---
title: "Notes"
author: "Eirini Zormpa"
date: "18/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## To do

- [] check that the code works with whatever dataset was created the previous day
- [] send participants the dataset that we will use?

## Schedule

- 9:00 Introduction - ask what this will be. Can we already check that everyone has the dataset for the data viz bit?
- 9:15 Introduction from me
- 9:30 Setup, Plotting with ggplot2, Building plots iteratively, Exercise 1, Boxplot, Exercise 2
- 10:30 Break
- 10:45 Barplots, Exercise 3, Adding labels and titles, Faceting, ggplot2 themes, Exercise 4, Customisation
- 11:45 Break

## Script and code


### Setup

Okay so let's start with loading all the packages that we will need.
As before, we need the tidyverse package, which as I mentioned, also includes the ggplot2 package.
We'll also need the here package, which makes it easier to work with relative paths.

```{r load-packaages}
library(tidyverse)
library(here)

# install.packages("viridisLite")
library(viridisLite)
```


And then we'll read in the data we created yesterday for the visualisation.

```{r read-data}
interviews_plotting <- read_csv(here("data", "interviews_plotting.csv"))
```

If you weren't able to create the dataset yesterday, you can use this code

```{r}
interviews <- read_csv(here("data", "SAFI_clean.csv"))

interviews_plotting_back_up <- interviews %>%
  separate_rows(items_owned, sep = ";") %>%
  replace_na(list(items_owned = "no_listed_items")) %>%
  mutate(items_owned_logical = TRUE) %>%
  pivot_wider(names_from = items_owned,
              values_from = items_owned_logical,
              values_fill = list(items_owned_logical = FALSE)) %>%
  separate_rows(months_lack_food, sep = ";") %>%
  mutate(months_lack_food_logical = TRUE) %>%
  pivot_wider(names_from = months_lack_food,
              values_from = months_lack_food_logical,
              values_fill = list(months_lack_food_logical = FALSE)) %>%
  mutate(number_months_lack_food = rowSums(select(., Jan:May))) %>%
  mutate(number_items = rowSums(select(., bicycle:car)))
```

### Plotting with ggplot2!

Anything that you put in the ggplot() function can be seen by any geom layers that you add, i.e. these are *universal* plot settings. This includes the x- and y-axis mappings.

Building plots with ggplot 2 is an iterative process. We start by defining the dataset, laying out the axes and other aesthetics we want to use, and choose a geom. Then we inspect the plot to see if it works. For example, here we should have 131 observations, each point being an observation. But I don't think we have 131 points!

```{r}
interviews_plotting %>%
  ggplot(aes(x = no_membrs, y = number_items)) +
  geom_point(alpha = 0.5)
```

```{r}
interviews_plotting %>%
  ggplot(aes(x = no_membrs, y = number_items)) +
  geom_jitter(height = 0.2, width = 0.2, colour = "blue")
```


```{r}
interviews_plotting %>%
  ggplot(aes(x = no_membrs, y = number_items, colour = village)) +
  geom_jitter(height = 0.2, width = 0.2)
```


```{r}
interviews_plotting %>%
  ggplot(aes(x = no_membrs, y = number_items, colour = village)) +
  geom_count()
```

## Boxplot

```{r}
interviews_plotting %>%
  ggplot(aes(x = respondent_wall_type, y = rooms)) +
  geom_jitter(height = 0.2, width = 0.2, colour = "tomato") +
  geom_boxplot(alpha = 0)
  
```

## Barplot

```{r}
library(viridisLite)

interviews_plotting %>%
  ggplot(aes(x = respondent_wall_type, fill = village)) +
  geom_bar(position = "dodge") +
  scale_fill_viridis_d()
```

```{r}
percent_wall_type <- interviews_plotting %>%
  count(village, respondent_wall_type) %>%
  group_by(village) %>%
  mutate(percent = n/sum(n) * 100) %>%
  ungroup()
```

- Make a point of removing cement
- Make a point of putting village in the x axis

```{r}
percent_wall_type %>%
  ggplot(aes(x = village, y = percent, fill = respondent_wall_type)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_viridis_d()
```

## Adding titles and labels

```{r}
percent_wall_type %>%
  ggplot(aes(x = village, y = percent, fill = respondent_wall_type)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_viridis_d() +
  labs(title = "Wall type by village",
       fill = "",
       y = "Percent",
       x = "")
```

```{r}
percent_wall_type %>%
  ggplot(aes(x = respondent_wall_type, y = percent)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_viridis_d() +
  labs(title = "Wall type by village",
       fill = "",
       y = "Percent",
       x = "") +
  facet_wrap(~village)
```

```{r}
percent_wall_type %>%
  ggplot(aes(x = village, y = percent, fill = respondent_wall_type)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_viridis_d() +
  labs(title = "Wall type by village",
       fill = "",
       y = "Percent",
       x = "") +
  theme_bw()
```